paper3

### 2022-07-29 paper3.Rmd
shap summary plot

rm(list = c("installed_packages", "packages"))
```
## Bayesian XXX by Evan Miller
Where zα/2 is the 1−α/2 quantile of a normal distribution.

@@

# split train/test set
# X = landmap_B.loc[:, ['Scenicness.mean', 'Wildness.mean']]
X = landmap_B.loc[:, ['Scenicness.median', 'x', 'y']]
y = landmap_B.loc[:, ['ScenicQuality']]

@@

# -- Create train/test split
from sklearn.model_selection import train_test_split
X_train, X_test, y_train, y_test = train_test_split(X, y, 
                                                    test_size = 0.2, 
                                                    random_state = 2046, 
                                                    stratify = y)
```
## Hyperparameter Tuning

@@

# Sample values from these configuration spaces using the routines in hyperopt.pyll.stochastic
from hyperopt.pyll.stochastic import sample
space = {'eta': hp.uniform('eta', 0.01, 1.01)}
print(sample(space))
```
#### 2. Define an objective/loss/cost function

@@

# Return the best parameters
space_eval(param_space, best_params)
```
#### 4. Results

@@

## Fit xgboost model

@@

# Consultant B
params = {
          'booster': 'gbtree',
        #'num_round': 1433.996,
              'eta': 0.892,
            'gamma': 0,
        'max_depth': 4,
 'min_child_weight': 0.6,
        'subsample': 0.500,
 'colsample_bytree': 1,
        'num_class': 4,
        'objective': 'multi:softmax',
      'eval_metric': 'merror',
             'seed': 1997
  }

@@

model = xgb.train(space_eval(param_space, best_params), 
                  dtrain, 
                  num_boost_round = 500,
                  evals = [(dtrain, 'train'), (dtest, 'test')],
                  early_stopping_rounds = 250)

# model = xgb.train(params, 
#                   dtrain, 
#                   num_boost_round = 500, 
#                   evals = [(dtrain, 'train'), (dtest, 'test')],
#                   early_stopping_rounds = 250)

@@

```{python}
import shap
import numpy as np
import matplotlib.pyplot as plt
from sklearn.preprocessing import StandardScaler
from sklearn.metrics import plot_confusion_matrix

# -- load the saved model
model = xgb.XGBRegressor()
model.load_model("Z:\\R\\R projects\\LCA-GB\\models\\model_B.json")

# -- view the confusion matrix
class_names = ['Low', 'Moderate', 'High', 'Outstanding']
disp = plot_confusion_matrix(model, X_test, y_test, 
                             display_labels = class_names, 
                             cmap = plt.cm.Blues, 
                             xticks_rotation = 'vertical')

# -- view the feature importance
importances = model.feature_importances_
indices = np.argsort(importances)
features = X.columns

@@

fig.set_size_inches(18.5, 10.5)
fig.savefig("Z:\\R\\R projects\\LCA-GB\\models\\pyplot.png")
plt.clf()

# -- play shap
explainer = shap.TreeExplainer(model)
shap_values = explainer.shap_values(X_test, approximate = True)
shap.summary_plot(shap_values[0], X_test.values, feature_names = X_test.columns)
fig.set_size_inches(18.5, 20.5)
fig.savefig("Z:\\R\\R projects\\LCA-GB\\models\\shap.png")

# -- first trial
shap.initjs()
shap_values = shap.TreeExplainer(model).shap_values(X_test)
shap.summary_plot(shap_values, X_test)

# -- second trial
scaler = StandardScaler()
scaled_data = scaler.fit_transform(X_test.iloc[:,:-1])
shap_values = explainer.shap_values(scaled_data.iloc[0,:-1].values.reshape(1,-1))
shap.force_plot(explainer.expected_value[0], shap_values[0], link='logit')  # repeat changing 0 for i in range(0, 5)

# -- third trial
shap.initjs()
class_names = ['Low', 'Moderate', ' High', 'Outstanding']
shap.summary_plot(shap_values, X_test, plot_type="bar", class_names= class_names, feature_names = ['Scenicness', 'Wildness'])

# -- plot the global importance of each feature
shap.plots.bar(explainer)

# visualize the first prediction's explanation
shap.force_plot(explainer.expected_value[0], shap_values[0])

# fig, ax = plt.subplots(figsize=(20, 20),dpi=200)
# shap.summary_plot(shap_values)
# # or 
# shap.plots.beeswarm(shap_values)
shap.summary_plot(shap_values[2], X_test)

# fig, ax = plt.subplots(figsize=(20, 20),dpi=200)
shap.summary_plot(shap_values, X_test, plot_type="bar")

shap_interaction_values = shap.TreeExplainer(model).shap_interaction_values(X_train)

X_names = ['Scenicness.median','x_coord','y_coord']

shap.summary_plot(shap_interaction_values, X_train, max_display=15, 
                  feature_names = X_names,
                  plot_type = 'dot')
                  # plot_type="compact_dot")

import shap  # package used to calculate Shap values

# Create object that can calculate shap values
explainer = shap.TreeExplainer(my_model)

# calculate shap values. This is what we will plot.
# Calculate shap_values for all of val_X rather than a single row, to have more data for plot.
shap_values = explainer.shap_values(val_X)

# Make plot. Index of [1] is explained in text below.
shap.summary_plot(shap_values[1], val_X)                  
```

>>>>

rm(list = c("installed_packages", "packages"))
```

## Bayesian XXX by Evan Miller

Where zα/2 is the 1−α/2 quantile of a normal distribution. - Evan Miller shows a Bayesian approach to ranking 5-star ratings: - <https://stackoverflow.com/questions/1411199/what-is-a-better-way-to-sort-by-a-5-star-rating/40958702#40958702>

@@

# split train/test set
# X = landmap_B.loc[:, ['Scenicness.mean', 'Wildness.mean']]
X = landmap_B.loc[:, ['Scenicness.median_of_bayescore', 'Wildness.median', 'x', 'y']]
y = landmap_B.loc[:, ['ScenicQuality']]

@@

# -- Create train/test split
from sklearn.model_selection import train_test_split
X_train, X_test, y_train, y_test = train_test_split(X, y, 
                                                    test_size = 0.2, 
                                                    random_state = 2046, 
                                                    stratify = y)
```

## Hyperparameter Tuning

@@

# Sample values from these configuration spaces using the routines in hyperopt.pyll.stochastic
from hyperopt.pyll.stochastic import sample
space = {'eta': hp.uniform('eta', 0.01, 1.01)}
print(sample(space))
```

#### 2. Define an objective/loss/cost function

@@

# Return the best parameters
space_eval(param_space, best_params)
```

#### 4. Results

@@

## Fit/Train the xgboost model

@@

# Consultant B
params = {
          'booster': 'gbtree',
        'num_round': 1434, #1433.996,
              'eta': 0.892,
            'gamma': 0,
        'max_depth': 4,
 'min_child_weight': 0.6,
        'subsample': 0.500,
 'colsample_bytree': 1,
        'num_class': 4,
        'objective': 'multi:softmax',
      'eval_metric': 'merror',
             'seed': 1995
  }

@@

# model = xgb.train(space_eval(param_space, best_params),
#                   dtrain,
#                   num_boost_round = 500,
#                   evals = [(dtrain, 'train'), (dtest, 'test')],
#                   early_stopping_rounds = 250)

model = xgb.train(params,
                  dtrain,
                  num_boost_round = 500,
                  evals = [(dtrain, 'train'), (dtest, 'test')],
                  verbose_eval = 100,
                  early_stopping_rounds = 251)

@@

```{python, engine.path = '.venv/Scripts/python'}
import shap
import numpy as np
import matplotlib.pyplot as plt
import matplotlib.pylab as pl
import xgboost as xgb
from sklearn.preprocessing import StandardScaler
from sklearn.metrics import plot_confusion_matrix

# print the JS visualization code to the notebook
shap.initjs()

# -- load the saved model
model = xgb.XGBClassifier()
model.load_model("Z:\\R\\R projects\\LCA-GB\\models\\model_B.json")
```

## Confusion matrix
```{python}
# -- view the confusion matrix
class_names = ['Low', 'Moderate', 'High', 'Outstanding']
disp = ConfusionMatrixDisplay.from_predictions(model, X_test, y_test, 
                             display_labels = class_names, 
                             cmap = plt.cm.Blues, 
                             xticks_rotation = 'vertical')
```

```{python}
# -- play shap
explainer = shap.TreeExplainer(model)
shap_values = explainer.shap_values(X_train, approximate=True)

shap.plots.beeswarm(shap_values, show=False, color_bar=False)
plt.colorbar()
plt.show()

# shap.summary_plot(shap_values[1], 
#                   X_train.values, 
#                   feature_names = X_train.columns,
#                   plot_size = (10, 4))
#                   
# plt.gcf().axes[-1].set_aspect('auto')
# plt.tight_layout()
# As mentioned, smaller "box_aspect" value to make colorbar thicker
# plt.gcf().axes[-1].set_box_aspect(200)

```

```{python}
# -- view the feature importance
importances = model.feature_importances_
indices = np.argsort(importances)
features = X.columns

@@

fig.set_size_inches(18.5, 10.5)
fig.savefig("Z:\\R\\R projects\\LCA-GB\\models\\pyplot.png")
plt.clf()
```

```{python}
# get just the explanations for the positive class
# shap_values = shap_values[...,1]

# shap.plots.bar(shap_values[...,1])
```

```{python}
shap.initjs()
# shap_values[1] what does the index mean?
shap.summary_plot(shap_values[1], X_train.values, feature_names=X_train.columns)
# shap.summary_plot(shap_values, X_test, max_display = X_test.shape[1])

# -- Dependence Contribution Plots
# make plot.
shap.dependence_plot('Scenicness.median_of_bayescore', shap_values[3], X_train, interaction_index = 'Wildness.median')

# explainer = shap.TreeExplainer(model2)
# shap_values = explainer.shap_values(X_sampled)
# shap.summary_plot (shap_values, X_sampled, max_display=X_sampled.shape[1]) 
# i=2
# shap.force_plot(explainer.expected_value[0], shap_values[0][i], X_test.values[i], feature_names = X_test.columns)
# 
# 
# fig.set_size_inches(18.5, 20.5)
# fig.savefig("Z:\\R\\R projects\\LCA-GB\\models\\shap.png")
# 
# # -- first trial
# shap.initjs()
# shap_values = shap.TreeExplainer(model).shap_values(X_test)
# shap.summary_plot(shap_values, X_test)
# 
# # -- second trial
# scaler = StandardScaler()
# scaled_data = scaler.fit_transform(X_test.iloc[:,:-1])
# shap_values = explainer.shap_values(scaled_data.iloc[0,:-1].values.reshape(1,-1))
# shap.force_plot(explainer.expected_value[0], shap_values[0], link='logit')  # repeat changing 0 for i in range(0, 5)
# 
# # -- third trial
# shap.initjs()
# class_names = ['Low', 'Moderate', ' High', 'Outstanding']
# shap.summary_plot(shap_values, X_test, plot_type="bar", class_names= class_names, feature_names = ['Scenicness', 'Wildness'])
# 
# # -- plot the global importance of each feature
# shap.plots.bar(explainer)
# 
# # visualize the first prediction's explanation
# shap.force_plot(explainer.expected_value[0], shap_values[0])
# 
# # fig, ax = plt.subplots(figsize=(20, 20),dpi=200)
# # shap.summary_plot(shap_values)
# # # or 
# # shap.plots.beeswarm(shap_values)
# shap.summary_plot(shap_values[2], X_test)
# 
# # fig, ax = plt.subplots(figsize=(20, 20),dpi=200)
# shap.summary_plot(shap_values, X_test, plot_type="bar")
# 
# shap_interaction_values = shap.TreeExplainer(model).shap_interaction_values(X_train)
#     
# X_names = ['Scenicness.median','x_coord','y_coord']
# 
# shap.summary_plot(shap_interaction_values, X_train, max_display=15, 
#                   feature_names = X_names,
#                   plot_type = 'dot')
#                   # plot_type="compact_dot")
#                   
# import shap  # package used to calculate Shap values
# 
# # Create object that can calculate shap values
# explainer = shap.TreeExplainer(my_model)
# 
# # calculate shap values. This is what we will plot.
# # Calculate shap_values for all of val_X rather than a single row, to have more data for plot.
# shap_values = explainer.shap_values(val_X)
# 
# # Make plot. Index of [1] is explained in text below.
# shap.summary_plot(shap_values[1], val_X)                  
```
